---
title: "Summarize Data by Species"
subtitle: "Has there been a change in the timing of flowering time in the tropics"
author: "Sky and Kim"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r, echo = FALSE, include = FALSE}
# Change identifiers to your system and file naming. 
user <- "C:/Users/skyla" # computer user name
project <- "/OneDrive/Desktop/ChapterOne"
chapter.fp <- paste0(user, project)

location <- "Udzungwa"
```

```{r Set CRAN, echo = FALSE, include = FALSE}
# Global settings
local({r <- getOption("repos")
       r["CRAN"] <- "http://cran.r-project.org"
       options(repos = r)})

# install.packages("tidyverse")
#install.packages("knitr")
#install.packages("kableExtra")
#install.packages("ggplot2")

library(tidyverse)
library(knitr)
library(kableExtra)
library(ggplot2)
```

```{r, echo = FALSE, include = FALSE}
# Set path names for folders

# Define folder for data
Data.fp <- file.path(paste0(chapter.fp, "/Data"))
if (!dir.exists(Data.fp)) dir.create(Data.fp)

#Define foulder for results
Results.fp <- file.path(paste0(chapter.fp, "/Results"))
if (!dir.exists(Results.fp)) dir.create(Results.fp)

list.files(chapter.fp)
```

```{r}
#### Load the data file

# Read in data file
data_csv <- read.csv(paste0(Data.fp, "/", location, ".csv"), 
                    header = TRUE)
colnames(data_csv)

data_csv <- data_csv %>%
  dplyr::rename("species" = "gbif.canonicalName")

```

```{r}
#### Load the data file

sp_summary <- data_csv %>% 
  group_by(species) %>%
  summarise(count = n()) %>%
  dplyr::arrange(desc(count))
sp_summary

```

```{r}
#### filter out less than 30 species

sp_sum_df_30 <- sp_summary %>%
  as.data.frame() %>%
  dplyr::filter(count >= 30)

num_30 <- length(rownames(sp_sum_df_30))

sp_sum_df_30

write.csv(x = sp_sum_df_30,
          file = paste0(Results.fp, "/", location, "_30_species.csv"))

list.files(Results.fp)
```

There are `r num_30` species greater than or equal to 30 in the `r location` data set.

```{r}
#### filter to 20 years

# Save just the species names
spp <- sp_sum_df_30$species
outcome_df$Species <- spp

outcome_df <- data.frame(matrix(nrow = length(spp), ncol = 2))
colnames(outcome_df) <- c("Species", "Keep_Drop")

for (i in 1:length(spp)) {
data_csv_filtered <- data_csv %>%
  dplyr::select(species, Year) %>%
  dplyr::filter(species == spp[[i]]) %>%
  dplyr::arrange(Year)

year_range <- max(data_csv_filtered$Year) - min(data_csv_filtered$Year)

if (year_range >= 20) {
  outcome <- "Keep"
} else {
  outcome <- "Drop"
}
outcome_df$Species[i] <- spp[i]
outcome_df$Keep_Drop[i] <- outcome

you_in <- outcome_df %>%
  dplyr::filter(Keep_Drop == "Keep") %>%
  dplyr::select(Species) 

you_in <- you_in$Species

data_csv_years <- data_csv %>%
  dplyr::filter(species %in% all_of(you_in))
}

write.csv(x = outcome_df,
          file = paste0(Results.fp, "/", location, "_data_csv_years.csv"))
list.files(Results.fp)

# This filters out the species without genera and species
empty <- data_csv %>%
  dplyr::filter(species == "")
```

### Outcome of setting a cut off of 20 years minimum for the dataset. 
`r kable(outcome_df) %>% kable_styling(full_width = F, position = "left")`

### Count for each species in the `r location`` data set.
`r kable(sp_sum_df_30) %>% kable_styling(full_width = F, position = "left")`

```{r, include = TRUE}
Species_list <- sp_sum_df_30$species

# Create empty data frame with 0 rows and 3 columns
df <- data.frame(matrix(ncol = 3, nrow = num_30))

# Provide column names
colnames(df) <- c("species", "pValue", "R2")

# Run a linear regression
for (i in 1:length(Species_list)) {
  df$species[[i]] <- Species_list[[i]]
  species_data <- data_csv %>% dplyr::filter(species == Species_list[i])
  sum <- summary(lm(Day.of.the.year ~ Year, data = species_data))
  df$pValue[[i]] <- round(sum$coefficients[8], 4)
  df$R2[[i]] <- round(sum$r.squared, 4)
}

#test
sum1<- lm(Day.of.the.year~Year*Species, data=species_data)
summary(sum1)
library(lmerTest)
lmer(Day.of.the.year~Year*Species+(1|Genus|family), data=species_data)
library(emmeans)
pairs(emmeans(sum1,~Species))
library(sjPlot)
plot_model(sum1,type="coeff")

# Sort the species by p value, lowest to highest
df <- df %>%
  dplyr::arrange(pValue)

write.csv(x = df,
          file = paste0(Results.fp, "/", location, "_LinearRegression_species.csv"))
```

### Linear regression of Flowering Day by Year
`r kable(df) %>% kable_styling(full_width = F, position = "left")`
```{r, include = TRUE}
#Figures of each location

library(ggplot2)

ggplot(species_data, aes(x = Year, y = Day.of.the.year)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "blue")
```

